\plaintitle

### 2017-01-08 drafts/labor/introduction.tex
abstract and intro

\topic{The past decade has seen a flourishing of \textit{on--demand work}
driven by the reformulation of work as
constituent parts of larger tasks.}
This framing of work into modular components
has enabled the computational hiring \& management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed workers then engage in work whenever their schedules allow,
often with little to no awareness of the broader context of the work, and
often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
In this paper, we use ``on--demand work'' to capture a pair of related phenomena:
first, \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
and second, \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, or administrative support.
The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
such as AMT to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems and designs that enable new forms of production
(e.g. \cite{bernsteinSoylent,vizwiz,paolacci2010running}).

\topic{As on--demand work has grown far beyond information work, it has given rise to scholarship exploring myriad facets of on--demand work.}
\citeauthor{howe2008crowdsourcing} first described crowdsourcing as
``outsourcing to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.
However, for years its instantiation was limited to the utilization of
human intelligence to process data, participate in scientific studies, and perform information work~\cite{CrowdsourcingUserStudies,movieSummarizationWu,yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
More recently, crowdsourcing of physically embodied work
--- driving and cleaning, for instance ---
has become a focus for on--demand labor markets~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
This growth prompted efforts to understand not just the work, but also the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
Some of this research has been motivated by the identification of the sociality of gig work~\cite{crowdcollab},
and the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
Other work has focused on the outcomes of this frustration,
reflecting on the resistance workers express against digitally mediated labor markets~\cite{uberAlgorithm,dynamo}.

\topic{This body of research has broadly sought to answer one central question:
What does the future hold for on--demand work and those who do it?}
Researchers have offered insights on this question along three major threads:
First, \nameref{sec:complexity}?
      Specifically,
      \begin{inlinelist}[label=(\alph*)]
        \item How complex are the goals that crowd work can accomplish?, and
        \item What kinds of goals and industries may eventually utilize it?
      \end{inlinelist}
      \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
            Nebeling:2016:WCW:2858036.2858169,
            Hahn:2016:KAB:2858036.2858364}.
Second, \nameref{sec:decomposition}
      \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
            LykourentzouPersonalityMatters,KucherbaevReLauncher,
            Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
            Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}.
And third, \nameref{sec:relationships}?
      \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
            whyWouldAnyoneBrewer,takingAHITMcInnis}.
% \end{numberlist}

@@

More concretely, by looking at on--demand work as
an instantiation (or even a continuation) of piecework,
and by interrogating patterns that the corresponding literature predicts
on this basis, we can
first, make sense of past events as part of a much larger series of an interrelated phenomenon;
second, reflect on differences in the factors that impacted piecework historically and their impact on--demand work today;
and third, to some extent, offer predictions of what on--demand work researchers,
and workers themselves,
might expect to see on the horizon.
For example, we will draw on the piecework literature regarding task decomposition,
which was historically limited by shortcomings in measurement and instrumentation, and
leverage that understanding to suggest how modern technology affects this mechanism in on--demand work
--- namely, by enabling precise tracking and measurement of workers via algorithms and software.
By doing so, on--demand labor will find ways to decompose tasks to unprecedented levels.

We organize this paper as follows:
first, we review the definition and history of piecework
to make clear the analogy to on--demand work
(which we'll call \textit{crowd work} hereon
for consistency with the existing literature);
second, we interrogate the three major research questions above using the lens of piecework. 
We will identify similarities and differences between piecework as historically understood and
on--demand work as we know it today;
third, we will make predictions of future developments based on how those similarities and differences influenced piecework;
finally, we will offer implications for researchers and practitioners based on our results.

>>>>

\topic{The past decade has seen a flourishing of computationally-mediated labor.}
The framing of work into modular components
has enabled computational hiring and management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed workers engage in work whenever their schedules allow,
often with little to no awareness of the broader context of the work, and
often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.

For years, such labor was limited to information work such as data annotation and surveys~\cite{CrowdsourcingUserStudies,movieSummarizationWu,yuenSurvey,geiger2011managing,quinnbedersonTaxonomy}.
Howver, physically embodied work such as driving and cleaning have now spawned multiple online labor markets as well~\cite{uberAlgorithm,uberOfficial,zaarlyOfficial,taskrabbitOfficial}.
In this paper we will use the term \textit{on--demand work} to capture this pair of related phenomena:
first, \textit{crowd work}~\cite{crowdworkFuture}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work;
and second, \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, and administrative support.
%\topic{As on--demand work has grown, it has given rise to scholarship exploring the domains it occurs in and the people who engage in it.}
%When \citeauthor{howe2008crowdsourcing} coined crowdsourcing, he defined it in relationship to paid labor:
%``taking a function once performed by employees and outsourcing to an undefined, generally large group of people in the form of an open call''~\cite{howe2008crowdsourcing}.

The realization that complex goals can be accomplished by directing  crowds of workers has spurred the technology industry to flock to sites of labor
such as AMT to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems that enable new forms of production
(e.g. \cite{bernsteinSoylent,vizwiz,paolacci2010running}) and pursuing social scientific inquiry into the workers on these platforms~\cite{Ross,whoareNOTtheTurkers}.
This research has called out identified the sociality of gig work~\cite{crowdcollab},
as well as the frustration and disenfranchisement that these systems embody~\cite{turkopticon,martin2014being,takingAHITMcInnis}.
Others have focused on the outcomes of this frustration,
reflecting on the resistance that workers express against digitally-mediated labor markets~\cite{uberAlgorithm,dynamo}.

\topic{This body of research has broadly sought to answer one central question:
What does the future hold for on--demand work and those who do it?}
Researchers have offered insights on this question along three major threads:
First, \namerefl{sec:complexity}~\cite{crowdForgeKittur}?
      Specifically, how complex are the goals that crowd work can accomplish, and what kinds of goals and industries may eventually utilize it?
      \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
            Nebeling:2016:WCW:2858036.2858169,
            Hahn:2016:KAB:2858036.2858364}.
Second, \namerefl{sec:decomposition}
      \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
            LykourentzouPersonalityMatters,KucherbaevReLauncher,
            Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
            Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}?
And third, \namerefl{sec:relationships}
      \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
            whyWouldAnyoneBrewer,takingAHITMcInnis}?
% \end{numberlist}

@@

More concretely, by looking at on--demand work as
an instantiation (or even a continuation) of piecework,
and by interrogating patterns that the historical literature identifies, we can achieve three goals.
First, we can make sense of past events as part of a much larger series of an interrelated phenomenon.
Second, we can reflect on differences in the features that impacted piecework historically and on--demand work today.
And third, to some extent, we can use these differences to offer some predictions of what on--demand work researchers,
and workers themselves,
might expect to see on the horizon.
For example, we will draw on the piecework literature regarding task decomposition,
which was historically limited by shortcomings in measurement and instrumentation, and
leverage that understanding to suggest how modern technology affects this mechanism in on--demand work
--- namely, by enabling precise tracking and measurement of workers via algorithms and software.

We organize this paper as follows:
first, we review the definition and history of piecework
to make clear the analogy to on--demand work;
second, we interrogate the three major research questions above using the lens of piecework. 
For each question, we will survey piecework's historical perspective as well as on--demand work's current perspective, call out similarities and differences, and then hazard predictions on what piecework's history might suggest about the answers to these persistent questions for on--demand work.
